Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach

### 2014-05-23 manuscripts/manuscript.Rmd
adding more marc edits and comments

## Background on the Gaussian Process

<!-- 
- Background on non-parametric modeling.  
-->

<!--
- Definition
- Previous application
-->

We use a Bayesian nonparameteric regression approach based on Gaussian
process priors.  The use of Gaussian process (GP) regression (or
"kriging" in the geospatial literature) to formulate a predictive
model is relatively new in the context of modeling dynamical systems
[@Kocijan2005], and was first introduced in the context ecological
modeling and fisheries management in @Munch2005.  GP models have
subsequently been used to test for the presence of Allee effects
[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
determine temporal variation in food availability [@Sigourney2012], and
provide a basis for identifying model-misspecification [@Thorson2014].
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006. The nonparametric moniker for the
GP approach is unfortunate as (a) it still involves the estimation of
parameters, and (b), statisticians also use non-parametric to mean models
that make only weak distributional assumptions  [e.g. @Lehmann1975].
However, we retain the terminology for consistency with previous studies.

@@

Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
'best-case' scenario for the parametric model performance relative to
the GP.   This a best-case scenario for the parametric models because the underlying dynamics are always simuated
from one the parametric models in the set, whereas in reality we never
have the "true" model among our candidates.  By choosing one-dimensional
models with few parameters, we limit the chance that poor performance
will be due to our inability to estimate parameters well, something
that becomes a more severe problem for higher-dimensional parametric
models. Further, the parametric models we consider are those most commonly
used in modeling stock-recruitment dynamics or to model sudden transitions
between alternative stable states.

@@

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows:

We consider three parametric models that relate population size at time
$t+1$, $X_{t+1}$, to population size at time $t$, $X_t$  and harvest at
time $t$, $h_t$ by

$$ X_{t+1}= Z_t f(S_t,p) $$

Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]

$$S_t = X_t - h_t$$

or the Ricker model,

@@

where $K$ is the carrying capacity and $r$ the growth rate.  For the Allen model,

(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} $$

Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable (tipping) point.  For the Myers model,

(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$

Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an Allee effect with alternative stable states exist at $\theta > 2$).

The familiar Ricker model involves two parameters, corresponding
to a growth rate and a carrying capacity, and cannot support alternative
stable growth rate dynamics. The Allen model resembles the Ricker
dynamics with an added Allee effect parameter [@Courchamp2008], below
which the population cannot persist.  The Myers model also has three
parameters and contains an Allee threshold, but has compensatory rather
than over-compensatory density dependence (resembling a Beverton-Holt
curve rather than a Ricker curve at high densities.)  
We assume multiplicative log-normal noise perturbs the growth predicted 
by the each of the deterministic model skeletons described above. This 
introduces one additional parameter $\sigma$ that must be estimated by each
model.

Because of our interest in management performance in the presence of
tipping points, all of our simulations are based on the Allen model.
The Allen model is thus structurally correct and is expected to provide
a best-case scenario when the 'true' dynamics are known.
The Ricker model is a reasonable approximation of these dynamics far from the Allee
threshold (but lacks threshold dynamics), while the Myers model shares
the essential feature of a threshold but differs in the structure.

We consider a period of `r Tobs` years of training data: long
enough that the estimates are not dependent on the particular realization,
while longer times are not likely to provide substantial improvement
(i.e. the results are not sensitive to this interval).  Each of the models
(described below) is fit to the same training data, as shown in Figure 1.

@@

We inferred posterior distributions for the parameters of each model in
a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R
[@RTeam] using jags, [@R2jags]).  We choose uninformative uniform priors
for all parameters (See appendix Tables S1-S3; R code provided). We
show one-step-ahead predictions of these model fits in Figure 1.
Each chain was tested for Gelman-Rubin convergence and results were
robust to longer runs.  For each simulation we also applied several
commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
to identify the best fitting model.

<!-- SDP via parametric models -->

An optimal policy function was then inferred through stochastic dynamic
programming for each model given the posterior distributions of the
parameter estimates.  This policy maximizes the expectation of the value
function integrated over the parameter uncertainty. (code implementing
this algorithm provided in the Appendix).

@@

@@

<!-- SDP via the model -->

@@

@@

```{r figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP. Points show the stock size simulated by the true model.  Overlay shows the range of states predicted by each model, based on the state observed in the previous time step. The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).  Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
y <- numeric(8)
y[1] <- 4.5
for(t in 1:(length(y)-1))
      y[t+1] = z_g() * f(y[t], h=0, p=p)
# predicts means, does not reflect uncertainty estimate!
crash_data <- step_ahead_posteriors(y)
crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
ggplot(crash_data) + 
  geom_boxplot(aes(as.factor(as.integer(time)), value, 
                   fill = variable, col=variable), 
               alpha=.7, outlier.size=1, position="identity") + 
#  geom_line(aes(time, value, col = variable, 
#            group=interaction(L1,variable)), alpha=.1) + 
  geom_point(aes(time, stock), size = 3) + 
  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
                      guide = guide_legend(override.aes = list(alpha = 1))) +  
  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
                      guide = guide_legend(override.aes = list(alpha = 1))) +  
  facet_wrap(~variable) + 
  theme(legend.position="none") + xlab("time") + ylab("stock size") 
```

@@

```{r figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
  facet_wrap(~method) + guides(legend.position="none")

@@

The consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated is shown in Figure 5.
As expected from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around it's un-fished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal profits.

The results across replicate stochastic simulations can most
easily be compared by using the relative differences in net present
value realized by each of the model. Figure 6 shows the distribution of values
(profits) realized by each simulation shown in Figure 5.  The GPDP
approach most consistently realizes a value close to the optimal solution,
and importantly avoids ever driving the system across the tipping point,
which results in the near-zero value cases in the parametric models.

@@

These results are not sensitive to the modeling details of the simulation.
The GPDP estimate remains very close to the optimal solution obtained
by knowing the true model across changes to the training simulation,
noise scale, parameters or structure of the underlying model, as seen in
Figure 7. The sensitivity analysis uses a factorial design of eight replicates 
of three different randomly generated parameter sets for each of two different 
generating models (Allen and Myers) over two different noise levels 
(0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
results are pooled across different random seeds (111, 222,
333), noise values of 0.01 and 0.05 (not large enough to violate the
self-sustaining criterion of @Reed1979), and different randomly generated
parameters sets for each model, and using either the Myers or Allen as the
underlying structure.  While the GPDP performs slightly closer to optimal
on the smaller noise levels, and closer on the Allen model then the Myers
model, the differences are small compared to the values observed in using
the wrong parametric model.  The GPDP solution consistently avoids either
failing to harvest at all or crashing the stock, which results in the very
low profits/values seen in parametric models in the previous figure. The
sensitivity analysis is compared against the optimal solution alone and
not shown against the estimates of the various parametric models (but
see Figure 6), where convergence of the Gibbs sampling could contribute
to their poorer performance if not checked by hand for each configuration.

Large noise relative to the distance between the stable and unstable
point, (or more precisely, between the optimal escapement and the unstable
point) increases the chance of a stochastic transition.  As a result,
the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be modeled
near-optimally. Models which underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters. Values near 1 indicate nearly optimal performance. Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=noise)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_wrap(~model, ncol=1)
```

@@

Though simple mechanistically motivated models offer the greatest
potential to increase our basic understanding of ecological processes
[@Cuddington2013; @Geritz2012], such models can be not only inaccurate but
misleading when relied upon in a quantitative decision making framework.
In this paper we have tackled two aspects of uncertainty that are both
common to many ecological decision-making problems and fundamentally
challenging to existing approaches which largely rely on parametric
models:

@@

### GPDP expresses larger uncertainty in regions where the data are poor

@@

Lastly, it should be noted that outside the data, the GP reverts to
the prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the
SDP.  In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its parameters
$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
In practice, both the choice of mean and the choice of the covariance
function may be chosen to confer particular biological properties,
as well as more biologically informed priors for $\ell$ and $\sigma$.
In principle, this may allow a manager to improve the performance of the
GPDP approach by adding only enough additional detail as is justified.
For instance, it would be possible to use a linear or a Ricker-shaped
mean in the prior without making the much stronger assumption that the
Ricker is the structurally correct model [@Sugeno2013a]. Future research
should focus on identifying criteria to ensure the prior and the value function
are chosen appropriately for the problem at hand.

>>>>

@@

Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
'best-case' scenario for the parametric model performance relative to
the GP.   This is abest-case scenario for the parametric models because the underlying dynamics are always simuated
from one the parametric models in the set, whereas in reality we never
have the "true" model among our candidates.  By choosing one-dimensional
models with few parameters, we limit the chance that poor performance
will be due to our inability to estimate parameters well, something
that becomes a more severe problem for higher-dimensional parametric
models. Further, the parametric models we consider are those most commonly
used in modeling stock-recruitment dynamics or to model sudden transitions
between alternative stable states.

@@

We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are

$$ X_{t+1}= Z_t f(X_t, h_t, p) $$

Where Z(t) is log-normally distributed process stochasticity [@Reed1979] and $p$ is a vector of parameters to be estimated from the data.   We describe three choices for $f(X_t,p)$ in the next section.In this simple model, the control action is a harvest or take, $h_t$, measured in the same units as $X$, at time
$t$.  Thus, in the presence of take, the population size on the right hand side of Eqn 1 is replaced by
$S_t=X_t-h_t$.

To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted by
$\Pi(x,h)$. For example, if the return is simply the harvest at time $t$, then $\Pi(x,h)=min(x,h)$. We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E_Z}[ \sum_{t = 0}^{T}  \Pi_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E_Z}$ denotes the expectation over the process stochasticity.

In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.  This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000, @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014].

\begin{equation}
V_t(x) = \max_{h_t} \left[\Pi(h_t, x_t) + \delta \cdot \mathbf{E_Z}  V_{t+1}( Z_{t+1} f(x - h_t)) \right]
\end{equation}

That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $\Pi(h_t, x_t)$ and the population size after the take is $x-h_t$, which is then translated into $X_{t+1}$ by Eqn 1. We consider the finite time problem with $T=$`r Tmax`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].

\subsubsection{Parametric models}\label{parametric-models}

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate. In the Ricker model,

@@

In the Allen model,

(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} $$

Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable steady state (i.e., the tipping point).

In the Myers model,

(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}$$

Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$ > 2 leads to Allee effects and multiple stable states.

The Ricker model involves two parameters, corresponding
to a growth rate and a carrying capacity, and cannot support alternative
stable growth rate dynamics. The Allen model resembles the Ricker
dynamics with an added Allee effect parameter [@Courchamp2008], below
which the population cannot persist.  The Myers model also has three
parameters and contains an Allee threshold, but has compensatory rather
than over-compensatory density dependence (resembling a Beverton-Holt
curve rather than a Ricker curve at high densities.)

The multiplicative log-normal stochasticity perturbs the growth predicted 
by each of these deterministic model skeletons. This 
introduces one additional parameter $\sigma$ that must be estimated.

Because of our interest in management performance in the presence of
tipping points, all of our simulations are based on the Allen model.
The Allen model is thus "structurally correct" and is expected to provide
a best-case scenario when the 'true' dynamics are known.
The Ricker model is a reasonable approximation of these dynamics far from the Allee
threshold (but lacks threshold dynamics), while the Myers model shares
the essential feature of a threshold but differs in the structure.

We consider a period of `r Tobs` years of training data: long
enough that the estimates do not depend on the particular realization,
while longer times are not likely to provide substantial improvement
(i.e. the results are not sensitive to this interval).  Each of the models
(described below) is fit to the same training data (Figure 1).

@@

We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
[@RTeam] using jags, [@R2jags]).  We choose uniform priors
for all parameters (See appendix Tables S1-S3; R code provided). We
show one-step-ahead predictions of these model fits in Figure 1.
We tested each chain for Gelman-Rubin convergence and results were
robust to longer runs.  For each simulation we also applied several
commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
to identify the best fitting model.

@@

 The use GP methods to formulate a predictive
model is relatively new in the context of modeling dynamical systems
[@Kocijan2005], and was first introduced in the context ecological
modeling and fisheries management in @Munch2005.  GP models have
subsequently been used to test for the presence of Allee effects
[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
determine temporal variation in food availability [@Sigourney2012], and
provide a basis for identifying model-misspecification [@Thorson2014].
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

The nonparametric moniker for the
GP approach is unfortunate as (a) it still involves the estimation of
parameters, and (b), statisticians also use non-parametric to mean models
that make only weak distributional assumptions  [e.g. @Lehmann1975].
However, we retain the terminology for consistency with previous studies.

<!-- FIXME Marc suggests putting the full GP formulation in here?? -->

@@

<!-- SDP via the model  FIXME much more detail here-->

The method of Gaussian Process Dynamic Programming
--------------------------------------------------

@@

Parametric and GP models for population dynamics
------------------------------------------------

@@

```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP. Points show the stock size simulated by the true model.  Overlay shows the range of states predicted by each model, based on the state observed in the previous time step. The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).  Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
y <- numeric(8)
y[1] <- 4.5
for(t in 1:(length(y)-1))
      y[t+1] = z_g() * f(y[t], h=0, p=p)
# predicts means, does not reflect uncertainty estimate!
crash_data <- step_ahead_posteriors(y)
crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
ggplot(crash_data) + 
  geom_boxplot(aes(as.factor(as.integer(time)), value, 
                   fill = variable, col=variable), 
               alpha=.7, outlier.size=1, position="identity") + 
#  geom_line(aes(time, value, col = variable, 
#            group=interaction(L1,variable)), alpha=.1) + 
  geom_point(aes(time, stock), size = 3) + 
  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
                      guide = guide_legend(override.aes = list(alpha = 1))) +  
  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
                      guide = guide_legend(override.aes = list(alpha = 1))) +  
  facet_wrap(~variable) + 
  theme(legend.position="none") + xlab("time") + ylab("stock size") 
```

@@

```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
  facet_wrap(~method) + guides(legend.position="none")

@@

The consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated is shown in Figure 5.
As expected from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around it's un-fished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal long-term catch.

The results across replicate stochastic simulations can most
easily be compared by using the relative differences in net present
value realized by each of the model. Figure 6 shows the distribution of values
realized by each simulation shown in Figure 5.  The GPDP
most consistently realizes a value close to the optimal solution,
and importantly avoids ever driving the system across the tipping point,
which results in the near-zero value cases in the parametric models.

@@

These results are not sensitive to the modeling details of the simulation.
The GPDP estimate remains very close to the optimal solution obtained
by knowing the true model across changes to the training simulation,
noise scale, parameters or structure of the underlying model, as seen in
Figure 7. The sensitivity analysis uses a factorial design of eight replicates 
of three different randomly generated parameter sets for each of two different 
generating models (Allen and Myers) over two different noise levels 
(0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
we pool results across different random seeds (111, 222,
333), noise values of 0.01 and 0.05 (not large enough to violate the
self-sustaining criterion of @Reed1979), and different randomly generated
parameters sets for each model, and using either the Myers or Allen as the
underlying structure.  While the GPDP performs slightly closer to optimal
on the smaller noise levels, and closer on the Allen model then the Myers
model, the differences are small compared to the values observed in using
the wrong parametric model.  The GPDP solution consistently avoids either
failing to harvest at all or crashing the stock, which results in the very
low values seen in parametric models in the previous figure. The
sensitivity analysis is compared against the optimal solution alone and
not shown against the estimates of the various parametric models (but
see Figure 6), where convergence of the Gibbs sampling could contribute
to their poorer performance if not checked by hand for each configuration.

Large noise relative to the distance between the stable and unstable
point, (or more precisely, between the optimal escapement and the unstable
point) increases the chance of a stochastic transition.  As a result,
the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be modeled
near-optimally. Models that underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

<!--FIXME the above para is not particularly convincing. Quantitative please? -->

```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values near 1 indicate nearly optimal performance. Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=noise)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_wrap(~model, ncol=1)
```

@@

Though simple mechanistically motivated models offer the greatest
potential to increase our basic understanding of ecological processes
[@Cuddington2013; @Geritz2012], such models can be both inaccurate and
misleading when relied upon in a quantitative decision making framework.
In this paper we have tackled two aspects of uncertainty that are both
common to many ecological decision-making problems and fundamentally
challenging to existing approaches which largely rely on parametric
models:

@@

### GPDP population dynamcis capture larger uncertainty in regions where the data are poor

<!-- FIXME More emphasis/explanation on _how_ it does this -->

@@

Lastly, it should be noted that outside the data, the GP reverts to
the prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the
SDP.  In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its parameters
$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
In practice, both the choice of mean and the choice of the covariance
function may be chosen to confer particular biological properties,
as well as more biologically informed priors for $\ell$ and $\sigma$.
In principle, this may allow a manager to improve the performance of the
GPDP by adding only enough additional detail as is justified.
For instance, it would be possible to use a linear or a Ricker-shaped
mean in the prior without making the much stronger assumption that the
Ricker is the structurally correct model [@Sugeno2013a]. Future research
should focus on identifying criteria to ensure the prior and the value function
are chosen appropriately for the problem at hand.
